---
title: "Assignment 3 - Empirical Findings"
author: "Md Mujahedul Islam & Johannes Schulz-Knappe"
date: "April 26, 2016"
output: pdf_document
subtitle: What explains the recent electoral success of the Alternative for Germany?
#bibliography: Library.bib
---


``` {r intialisation, echo = F, message = F, warning = F}
options(repos=c(CRAN="http://mirrors.softliste.de/cran/"))

rm(list = ls())

load <- c("magrittr", "dplyr", "xtable", "texreg", 
          "stargazer", "readxl", "ggplot2", "grid", 
          "captioner", "foreign", "rms", "Zelig", 
          "knitr", "GGally", "gridExtra", 
          "zoo", "rugarch", "texreg", "dplyr", "magrittr")

loaded <- lapply(load, function(x) {
  if(!require(x, character.only = T)) {
    install.packages(x)
    require(x, character.only = T)
  }
})
rm(load, loaded)

figcap <- captioner(prefix = "Figure", 
                    auto_space = TRUE, levels = 1, type = NULL)
tablecap <-captioner(prefix = "Table", 
                     auto_space = TRUE, levels = 1, type = NULL)

setwd("C:/Users/User/Documents/GitHub/Assignment03")
```


View(Data)
# Introduction

# Data and Variables
```{r echo=F, warning=F, message=F, error=F}
load(".RData")
```

## Data gathering



## Data cleaning and merging

# Empirical Findings


## Descriptive Statistics
## Summary Statistics
Please give a short description of this summary statistics!!!
```{r sumstat, echo=F, warning=F, message=F, error=F, results='asis'}
summary_labels <- c('Vote share of AfD', 'Vote turnout in 2011', 
                        'Vote share of CDU in 2011', 'Vote share of Greens in 2011', 
                        'Vote share of SPD in 2011', 'Vote share of FDP in 2011', 
                        'Vote share of Linke in 2011', 'High school ratio', 
                        'No high school degree', 'GDP per capita', 
                        'Unemployment rate', 'No. of refugees')


stargazer(Data[4:16],
title = 'Summary statistics of all variables',
covariate.labels = summary_labels,
font.size = 'small',
type="latex", header = F)
```

```{r histo, echo=F, warning=F, message=F, error=F, results='hide', fig.height=4}
ggplot(Data) +
  geom_histogram(aes(x = n.refugees),
                 colour = "black",
                 fill = "transparent",
                 binwidth = 30) +
  ylab("Frequency") +
  xlab("No. of refugess in RP, BW, and SA") +
  theme_light()
```


```{r bar, echo=F, warning=F, message=F, error=F, results='hide', fig.width=13}
ggplot(Data) +
  geom_bar(aes(y = vote.AfD, x = reorder(district.name, -vote.AfD)), 
           stat = "identity") +
  ylab("AfD's vote share") +
  xlab("Different consistuencies of RP, BW and SA") +
  ggtitle("Vote share of AfD in three different states of Germany") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90))
  ```





\pagebreak

## Correlations among variables

Please give a short description of this correlation matrix!!!


```{r corr, echo=F, warning=F, message=F, error=F, results='asis'}
 correlation.matrix <- cor(Data[,c("vote.AfD","lag.CDU",
                                    "lag.SPD","GDP.capita",
                                    "unempl.rate","n.refugees",
                                    "abitur.ratio")])
  stargazer(correlation.matrix, 
            title = "Correlation Matrix of different variables",
            font.size = 'small',
            type="latex", header = F)
```       
  

### Correlation of AfD's vote share and GDP per capita
The following figure shows statistically significant (the t value from the Pearson's product-moment correlation test is 6.27) positve correlation of increase in unemployment rate with the increase in AfD's vote share in those three states. 

``` {r fig1, echo = F, message = F, results = "hide", error=F, fig.height=4}
ggplot2::ggplot(Data,aes(log(x = GDP.capita), vote.AfD)) + 
  xlab("GDP per capita in RP, BW, and SA") +
  ylab("AfD's vote share (in %)") +
  geom_point() + geom_smooth() + 
  theme_bw()
```


\pagebreak

### Correlation of AfD's vote share and unemployment rate
The following figure shows statistically significant (t value is -2.04) negative coorelation of increasing Per capita GDP with AfD's vote share. The 

``` {r fig2, echo = F, message = F, results = "hide", error=F, fig.height=4}
ggplot2::ggplot(Data,aes(log(x = unempl.rate), vote.AfD)) + 
  xlab("Unemployment rate in RP, BW, and SA") +
  ylab("AfD's vote share (in %)") +
  geom_point() + geom_smooth() + 
  theme_bw()
```

\pagebreak

### Correlation of AfD's vote share and number of refugees

The result though shows positive coorelation of increased number of refugees with with AfD's vote share, it is not statistically significant since the t value is .13 from the Pearson's product-moment correlation test. 

``` {r fig3, echo = F, message = F, results = "hide", error=F, fig.height=4}
ggplot2::ggplot(Data,aes(log(x = n.refugees), vote.AfD)) + 
  xlab("Number of refugees in RP, BW, and SA") +
  ylab("AfD's vote share (in %)") +
  geom_point() + geom_smooth() + 
  theme_light()
```





\pagebreak

## Inferrential Statistics

``` {r regout, echo = F, message = F, results = "asis", error=F}
reg1 <- lm(vote.AfD ~ GDP.capita + unempl.rate + n.refugees, data = Data)
reg2 <- lm(vote.AfD ~ GDP.capita + unempl.rate + n.refugees + lag.CDU + 
             lag.SPD + lag.turnout, data = Data)
reg3 <- lm(vote.AfD ~ GDP.capita + unempl.rate + n.refugees + 
             abitur.ratio + nodegree.ratio, data = Data)
reg4 <- lm(vote.AfD ~ GDP.capita + unempl.rate + n.refugees 
           + lag.CDU + lag.SPD + lag.turnout + abitur.ratio + 
             nodegree.ratio, data = Data)


var_labels <- c('GDP per capita', 'Unemployment rate', 'No. of Refugees', 
            'Vote share of CDU in 2011 election',
            'Vote share of SPD in 2011 election',
            'High School Ratio', 'University Degree ratio', '(Intercept)')

stargazer::stargazer(reg1, reg2, reg3, reg4,
          omit = 'as.factor*', 
          omit.stat = c('f', 'ser'), # to nicely fits on the page
          out.header = F,
          title = 'Determinants of vote for the Alternative for Germany',
          dep.var.labels = 'Vote share of AfD',
          covariate.labels = var_labels,
          label = 'AfD_voteshare',
          add.lines = list(c('District FE?', rep('NO', 4))),
          font.size = 'small',
          type="latex", header = F)
```




